---
title: "Getting other data structures from iscream"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting other data structures from iscream}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
#bibliography: refs.bib
link-citations: yes
---

The examples here show how iscream output can be converted into other data
structures for further analysis.

```{r}
library(iscream)
data_dir <- system.file("extdata", package = "iscream")
bedfiles <- list.files(data_dir, pattern = "[a|b|c|d].bed.gz$", full.names = TRUE)
regions <- c(A = "chr1:1-6", B = "chr1:7-10", C = "chr1:11-14")
```

## GRanges

```{r, message=FALSE}
library(GenomicRanges)
```

GRanges can be used as the input to all of iscream's querying functions. The
output of `tabix()` and `make_mat()`, `make_bsseq_mat()` can also be turned into GRanges.

### From `tabix` queries

```{r}
tabix(bedfiles[1], regions) |> GRanges()
```

If the input is already a `GRanges` object, `tabix()` will also return a
`GRanges` object along with any attached metadata. With multiple input files, it
returns a `GRangesList`. `tabix()` uses `findOverlaps()` to join input metadata
with the queried data.

```{r}
gr <- GRanges(regions)
values(gr) <- DataFrame(
    gene = c("gene1", "gene2", "gene3"),
    some_metadata = c("s1", "s2", "s3")
)
gr
tabix(bedfiles, gr)
```

If the input BED file is not zero-based (e.g. Bismark coverage files), set
`zero_based = FALSE` in the `tabix()` call to get the correct conversion from
data frame to GenomicRanges.

### From `summarize_regions`

With a named set of regions, you'll need to `set_region_rownames` as the input
to create the GRanges object:

```{r}
(summary <- summarize_meth_regions(
  bedfiles,
  regions,
  set_region_rownames = TRUE,
  fun = c("sum", "mean"))
)
GRanges(rownames(summary), summary = summary)
```

If the regions vector is not named, pass the `feature` column as the GRanges
input regions:

```{r}
(summary <- summarize_meth_regions(
  bedfiles,
  unname(regions),
  fun = c("sum", "mean"))
)
GRanges(summary$feature, summary = summary[, -1])
```


## Making SummarizedExperiment objects

```{r, message=FALSE}
library("SummarizedExperiment")
```

### From `make_mat()` or `make_bsseq_mat()` matrices

```{r}
(mats <- make_bsseq_mat(bedfiles, regions))
(mats.sparse <- make_bsseq_mat(bedfiles, regions, sparse = TRUE))
gr <- GRanges(mats$chr, mats$pos)

# dense
SummarizedExperiment(assays = list(M = mats$M, Cov = mats$Cov), rowRanges = gr)

# sparse
SummarizedExperiment(
  assays = list(M = mats.sparse$M, Cov = mats.sparse$Cov),
  rowRanges = gr
)

```

## Making `BSseq` objects

A BSseq object is a type of SummarizedExperiment, but it cannot handle sparse
matrices:

```{r, eval=FALSE}
library(bsseq)
do.call(BSseq, mats)
```

## Session info

```{r}
sessionInfo()
```
